---
title: "Calculating heritability from simulation output"
author: "Kevin Thornton"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating heritability from simulation output}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Heritability in the entire population

Ingredients:

1. index.txt is the index file corresponding to the simulation output
2. phenotypes.bin.gz is the file containing the phenotypes of the individuals

To get the distribution of heritabilities in the entire population:

```{r}
idx=read.table(system.file("extdata", "index.txt", package = "diseaseSims"))
##open in binary mode...
f=gzfile(system.file("extdata", "phenotypes.bin.gz", package = "diseaseSims"),"rb")

h2=array()
for(i in 1:nrow(idx))
{
	N = readBin(f,"integer",1)
	m = matrix(readBin(f,"numeric",2*N),ncol=2,byrow=TRUE)
	##column 1 = genetic component, column2 = random component
	h2[i]=var(m[,1])/( var(m[,1]) + var(m[,2]) )
}
##Now, you can get the mean, etc., which will not be very accurate for this example,
##as it is based on just 3 replicates
print(mean(h2))
print(h2)
```

Another method involves calling the getPheno function to read the file.  This method is slower, as it requires constant opening of the file and seeking to the desired position, but it is handy for accessing specific records directly: 

```{r}
library(diseaseSims)
idx=read.table(system.file("extdata", "index.txt", package = "diseaseSims"))
pfile=system.file("extdata", "phenotypes.bin.gz", package = "diseaseSims")

h2=array()
OFFSET=0
for( i in 1:nrow(idx) )
{
	m = getPheno(pfile,OFFSET)
	h2[i]=var(m[,1])/( var(m[,1]) + var(m[,2]) )
	##Update the OFFSET
	OFFSET = OFFSET + idx$V3[i]
}
##Now, you can get the mean, etc., which will not be very accurate for this example,
##as it is based on just 3 replicates
print(mean(h2))
print(h2)
```